---
title: "BDSkyline"
author: "Angela"
date: "20/12/2019"
output: html_document
---

Code club 2020-01-22

Purpose of this script:
-read in multiple output log files from birth death skyline runs from BEAST
-visualize the reproduction number over time

Challenge:
-writing and executing functions in which global objects are defined

Source of example logfiles and additional info:
https://taming-the-beast.org/tutorials/Skyline-plots/

Louis de Plessis' bdskytools package:
https://github.com/laduplessis/bdskytools

```{r Run this chunk if you need to install these libraries, echo=FALSE}
# install.packages("devtools")
# library(devtools)
# devtools::install_github("laduplessis/bdskytools")
# install.packages("ggplot2")
# install.packages("dplyr")
# install.packages("tidyr")
# install.packages("wesanderson") #very important
# install.packages("RColorBrewer")
```

```{r setup libraries, results="hide"}
library(bdskytools)
library(ggplot2)
library(dplyr)
library(tidyr)
library(wesanderson)
library(RColorBrewer)
```

```{r Read in a single logfile and plot Ro over time}
#Read in the logfile
lf<-readLogfile("BEAST/hcv_bdsky_30M.log", burnin=0.1)

#extract the HPDs 
Re_sky <- getSkylineSubset(lf, "reproductiveNumber")
Re_hpd <- getMatrixHPD(Re_sky)
delta_hpd <- getHPD(lf$becomeUninfectiousRate_BDSKY_Contemp) #use getHPD for one dimensional parameter
s_hpd <- getHPD(lf$rho_BDSKY_Contemp) #use getHPD for one dimensional parameter

#Plot the stepwise Re
plotSkyline(1:10, Re_hpd, type='step', ylab="R") #note that 1:10 reflects the dimensions of Ro

#In order to plot the smooth skyline we have to marginalise our Re estimates 
#on a regular timegrid and calculate the HPD at each gridpoint. 
#It is usually a good idea to use a grid with more cells than the dimension of Re
#To do this we first calculate the marginal posterior at every time of interest 
#using the function gridSkyline and then calculate the HPD for each of the finer time intervals.
#The times to grid the skyline on (timegrid), refers to years in the past.
timegrid <- seq(0,400,length.out=101)
times <- 1993-timegrid #sequences were sampled in 1993
Re_gridded <- gridSkyline(Re_sky, lf$origin, timegrid)
Re_gridded_hpd <- getMatrixHPD(Re_gridded)

#setup colors
fav<-wes_palette("Zissou1")

#plot a smooth skyline
png("OutputFigures/Re_plot.png", width = 5, height = 5, units = "in", res=1000)
plotSkylinePretty(times, Re_gridded_hpd, type='smooth', axispadding=0.0, 
                  col=fav[1], fill=fav[2], 
                  xlab="Time", ylab=expression("R"[e]), side=2, yline=2.5, xline=2, xgrid=TRUE, 
                  ygrid=TRUE, gridcol=pal.dark(cgray), ylims=c(0,3.5))
while (!is.null(dev.list()))  dev.off()
```

This is all fine and well for 1 log file, but what about when you want to look at lots of log files, save many objects from those log files, and then compare across runs?

```{r Read in multiple logfiles and plot Ro over time for all - the inefficient way}
####LOGFILE1####

#Read in the logfile
lf1<-readLogfile("BEAST/hcv1.log", burnin=0.1)

#extract the HPD for reproduction number
Re_sky1 <- getSkylineSubset(lf1, "reproductiveNumber")
Re_hpd1 <- getMatrixHPD(Re_sky1)

#setup for plot
timegrid1 <- seq(0,400,length.out=101)
times1 <- 1993-timegrid1 #sequences were sampled in 1993
Re_gridded1 <- gridSkyline(Re_sky1, lf1$origin, timegrid1)
Re_gridded_hpd1 <- getMatrixHPD(Re_gridded1)

#plot a smooth skyline
png("OutputFigures/Re_plot1.png", width = 5, height = 5, units = "in", res=1000)
plotSkylinePretty(times1, Re_gridded_hpd1, type='smooth', axispadding=0.0, 
                  col=fav[1], fill=fav[2], 
                  xlab="Time", ylab=expression("R"[e]), side=2, yline=2.5, xline=2, xgrid=TRUE, 
                  ygrid=TRUE, gridcol=pal.dark(cgray), ylims=c(0,3.5))
while (!is.null(dev.list()))  dev.off()

####LOGFILE2####

#Read in the logfile
lf2<-readLogfile("BEAST/hcv2.log", burnin=0.1)

#extract the HPD for reproduction number
Re_sky2 <- getSkylineSubset(lf2, "reproductiveNumber")
Re_hpd2 <- getMatrixHPD(Re_sky2)

#setup for plot
timegrid2 <- seq(0,400,length.out=101)
times2 <- 1993-timegrid2 #sequences were sampled in 1993
Re_gridded2 <- gridSkyline(Re_sky2, lf2$origin, timegrid2)
Re_gridded_hpd2 <- getMatrixHPD(Re_gridded2)

#plot a smooth skyline
png("OutputFigures/Re_plot2.png", width = 5, height = 5, units = "in", res=1000)
plotSkylinePretty(times2, Re_gridded_hpd2, type='smooth', axispadding=0.0, 
                  col=fav[1], fill=fav[2], 
                  xlab="Time", ylab=expression("R"[e]), side=2, yline=2.5, xline=2, xgrid=TRUE, 
                  ygrid=TRUE, gridcol=pal.dark(cgray), ylims=c(0,3.5))
while (!is.null(dev.list()))  dev.off()

#### Try to plot them together ####
fav2 <- adjustcolor(fav, alpha.f = 0.3)

png("OutputFigures/Re_plot3.png", width = 5, height = 5, units = "in", res=1000)
par(mar=c(5,4,4,4)+0.1)
plotSkylinePretty(times1, Re_gridded_hpd1, type='smooth', axispadding=0.0, 
                  col=fav[1], fill=fav2[2], 
                  xlab="Time", ylab=expression("R"[e]), side=2, yline=2.5, xline=2, xgrid=TRUE, 
                  ygrid=TRUE, gridcol=pal.dark(cgray), ylims=c(0,4),xaxis=FALSE)

plotSkylinePretty(times1, Re_gridded_hpd2, type='smooth', axispadding=0.0, 
                  col=fav[5], fill=fav2[5], 
                  xlab="Time", ylab=expression("R"[e]), side=2, yline=2.5, xline=2, xgrid=TRUE, 
                  ygrid=TRUE, gridcol=pal.dark(cgray), ylims=c(0,4), new=TRUE, add=TRUE)
while (!is.null(dev.list()))  dev.off()

```

Issue: very annoying/ugly script to have to change object names for each log file...

Challenge: can you write a function that can be applied to a logfile to define all these objects (with unique names dependent on the input to the function) to feed into the figure?

```{r CHALLENGE 1: INSERT NEW SCRIPT HERE}
# define a vector of log file inputs
my_logfiles<-list("BEAST/hcv1.log","BEAST/hcv2.log")
#my_logfiles<-as.list(dir("BEAST","log",full.names=T))
names(my_logfiles)<-c("hcv1","hcv2")

# write a function that operates over a list of logfiles
lg_to_df <- function(logfile){
  
  #Read in the logfile
  lf<-readLogfile(logfile, burnin=0.1)

  #extract the HPD for reproduction number
  Re_sky <- getSkylineSubset(lf, "reproductiveNumber")
  Re_hpd <- getMatrixHPD(Re_sky)
  
  #setup for plot
  timegrid <- seq(0,400,length.out=101)
  times <- 1993-timegrid #sequences were sampled in 1993
  Re_gridded <- gridSkyline(Re_sky, lf$origin, timegrid)
  Re_gridded_hpd <- getMatrixHPD(Re_gridded)
  
  # make a list of the desired outputs
  our_list <- list(Re_gridded_hpd,times)
}

# Run this function over our list of log files
# Return a list of the desired outputs
log_out <- lapply(my_logfiles, lg_to_df)

# Make a plot using the dataframes within the output list
# Plot over a list
fav2 <- adjustcolor(fav, alpha.f = 0.3)

png("OutputFigures/Re_plot_improved_list_b.png", width = 5, height = 5, units = "in", res=1000)
par(mar=c(5,4,4,4)+0.1)

for (i in 1:(length(log_out))){
  plotSkylinePretty(log_out[[i]][[2]], log_out[[i]][[1]], type='smooth', axispadding=0.0, 
                    col=fav[i], fill=fav2[i], xlab="Time", ylab=expression
                    ("R"[e]), side=2, yline=2.5, xline=2, xgrid=TRUE, ygrid=TRUE,
                    gridcol=pal.dark(cgray), ylims=c(0,4), add=(i>1))
  }
while (!is.null(dev.list()))  dev.off()
```

If you have time...apply same strategy to improve code below

```{r plot HPD sampling proportion, become uninfectious, origin}
####Extract mean estimate (point) with 95% HPD interval for parameters####
#Calculate 95%HPD for nexus 1
delta_hpd1<-getHPD(lf1$becomeUninfectiousRate_BDSKY_Contemp) 
s_hpd1<-getHPD(lf1$rho_BDSKY_Contemp) 
or_hpd1<-getHPD(lf1$origin_BDSKY_Contemp) 
HPDs_1<-data.frame(delta_hpd1, s_hpd1, or_hpd1)
colnames(HPDs_1)<-c("BecomeUninfectiousRate","SamplingProportion","Origin")
rownames(HPDs_1)<-c("LowerHPD","Mean","UpperHPD")
HPDs_1$Estimate<-rownames(HPDs_1)
HPDs_1$Cluster<-as.factor("1")
HPDs_1<-HPDs_1%>%mutate(PeriodOfInfectiousness=1/BecomeUninfectiousRate) #change to its more interpretable reciprocal
HPDs_1<-HPDs_1%>%mutate(OriginYear=2018-Origin) #change year instead of age
HPDs_1<-HPDs_1 %>% select(-c("Origin","BecomeUninfectiousRate"))
HPDs_1.2<-HPDs_1 %>% tidyr::gather("Parameter","Value","PeriodOfInfectiousness","SamplingProportion","OriginYear")
HPDs_1.3<-HPDs_1.2 %>% tidyr::spread("Estimate","Value")

#Calculate 95%HPD for nexus 2
delta_hpd2<-getHPD(lf2$becomeUninfectiousRate_BDSKY_Contemp) 
s_hpd2<-getHPD(lf2$rho_BDSKY_Contemp) 
or_hpd2<-getHPD(lf2$origin_BDSKY_Contemp) 
HPDs_2<-data.frame(delta_hpd2, s_hpd2, or_hpd2)
colnames(HPDs_2)<-c("BecomeUninfectiousRate","SamplingProportion","Origin")
rownames(HPDs_2)<-c("LowerHPD","Mean","UpperHPD")
HPDs_2$Estimate<-rownames(HPDs_2)
HPDs_2$Cluster<-as.factor("2")
HPDs_2<-HPDs_2%>%mutate(PeriodOfInfectiousness=1/BecomeUninfectiousRate) #change to its more interpretable reciprocal
HPDs_2<-HPDs_2%>%mutate(OriginYear=2018-Origin) #change year instead of age
HPDs_2<-HPDs_2 %>% select(-c("Origin","BecomeUninfectiousRate"))
HPDs_2.2<-HPDs_2 %>% tidyr::gather("Parameter","Value","PeriodOfInfectiousness","SamplingProportion","OriginYear")
HPDs_2.3<-HPDs_2.2 %>% tidyr::spread("Estimate","Value")

#and merge into a single data frame
allHPDs<-rbind(HPDs_1.3, HPDs_2.3)

#Okay, now plot as faceted ggplot for each of the Parameters with point for mean and bar for interval
colz<-c(fav[1],fav[5])
ggplot(allHPDs)+
  geom_pointrange(aes(x=Cluster,y=Mean, ymin=LowerHPD, ymax=UpperHPD, color=Cluster))+
  scale_color_manual(values=colz)+
  facet_wrap(. ~Parameter,scales = "free")+
  labs(x=NULL, y="Mean (95% HPD)")+
  theme_bw()+
  theme()

ggsave("OutputFigures/Facetted_estimates.png", width = 7, height = 5)
```

```{r CHALLENGE 2: Insert improved code here!}

```

